---
title: "Practical Machine Learning Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First we open the csv. 

```{r echo=TRUE}
training <- read.csv('/Users/ronthovsky09/Desktop/coursera/notes_r/practical_machine_learning/pml-training.csv', na.strings = c("NA","NaN","","#DIV/0!"))
testing <-read.csv('/Users/ronthovsky09/Desktop/coursera/notes_r/practical_machine_learning/pml-testing.csv', na.strings = c("NA","NaN","","#DIV/0!"))
head(data)
```
Looking at the data set, we must process the data. We must split the training set into training and test sets. 
```{r echo=TRUE}
library(caret)

inTrain <- createDataPartition(training$classe, p=0.8, list=FALSE)
train <- training[inTrain,]
test <- testing[-inTrain,]
```

Next, we must exclude variables that has either no variances, or a lot of NAs. 

```{r echo=TRUE}
NZV <- nearZeroVar(train)
train <- train[,-NZV]
test <- test[,-NZV]

is_na <- apply(train, 2, function (x) mean(is.na(x))) > .95
train <- train[, -which(is_na, is_na==FALSE)]
test<- test[, -which(is_na, is_na==FALSE)]

train <- train[,-(1:5)]
test <- test[,-(1:5)]
```

```{r echo =TRUE}
#We have reduced the number of training variables from 160 to 54. 

names(train)
```

Let's take at the correlation heat map

```{r echo=TRUE}
library(corrplot)
corrMap<-cor(train[,-54])
corrplot(corrMap, method = 'color', type ='lower', tl.cex = 0.8, tl.col = rgb(0,0,0))
```
Above shows the correlation heat map. The darker the color, the higher the correlation. Last, we will select the models used for the dataset. 

```{r echo=TRUE}
#Random Forest Tree

fit_RF <- train(classe~., data=train, method='rf')
fit_RF$finalModel

predict_RF <- predict(fit_RF, test)
#accuracy: 99%
confusionMatrix(fit_RF, test$classe)

```
Generalized Boosted Model. 

```{r echo=TRUE}
library(caret)
fit_gbm <- train(classe~., data=train, method='gbm')
fit_gbm$finalModel

predict_gbm <- predict(fit_gbm, test)
#accuracy: 98.48%
confusionMatrix(fit_gbm, test$classe)

```

Proceed with random forest model. 

``` {r echo=TRUE}
predict(fit_RF,testing)

```


